---
title: "RepResPA1"
output: html_document
---

##Loading and preprocessing the data

```{r, message=FALSE}
library(lubridate)
library(dplyr)
```

```{r}
#Load file data previously downloaded
load_data <- function() {
  activity <- read.csv('activity.csv')
  activity$date <- ymd(activity$date)
  return(tbl_df(activity))
}

activity <- load_data()
head(activity)
```



##What is mean total number of steps taken per day?

```{r}
#Grouping data by day
by_day <- function(activity) {
  gro_by_day <- group_by(activity, date)
  summary_tbl <- summarise(gro_by_day,
                           total = sum(steps, na.rm = TRUE),
                           average = mean(steps, na.rm = TRUE),
                           median = median(steps, na.rm = TRUE))
  return(summary_tbl)
}
day_steps <- by_day(activity)
hist(day_steps$total, main="Distribution of total daily steps",
     xlab="Counts", ylab="Total Steps")
cat('Median number of the total number of steps taken per day:', median(day_steps$total, na.rm=TRUE))
cat('Mean number of the total number of steps taken per day:', mean(day_steps$total, na.rm=TRUE))

```



##What is the average daily activity pattern?

```{r}
#Grouped by 5 minute interval
step_interval <- function(activity) {
  gro_by_int <- group_by(activity, interval)
  summary_tbl <- summarise(gro_by_int,
                           total = sum(steps, na.rm = TRUE),
                           average = mean(steps, na.rm = TRUE))
  return(summary_tbl)
}
int_sum <- step_interval(activity)
plot(int_sum$interval, int_sum$total, type="l",
     main="Average number of steps taken across all days",
     xlab="Interval",
     ylab="Steps")

```

##Imputing missing values

There are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

```{r}
cat('NA:', sum(is.na(activity$steps)))

imp_mis <- function(activity, int_sum) {
  imp_act <- left_join(activity, int_sum, by='interval')
  imp_act$imp_ste <- ifelse(is.na(imp_act$steps), 
                                    imp_act$average, 
                                    imp_act$steps)
  return(select(imp_act, date, interval, steps, imp_ste)) 
}

imp_act <- imp_mis(activity, int_sum)
imp_act_daily_totals <- imp_act %>% 
                           group_by(date) %>% 
                           summarise(daily_total = sum(imp_ste))
hist(imp_act_daily_totals$daily_total,
     main="Imputed Totals",
     xlab="Imputed Steps",
     ylab="Count")
cat('Median total number of steps taken per day:',
    median(imp_act_daily_totals$daily_total, na.rm=TRUE))
cat('Mean total number of steps taken per days:',
    mean(imp_act_daily_totals$daily_total, na.rm=TRUE))
```

Yes, the values differ from the first part of the assignment.The impact is that the values have shifted to the right.



##Are there differences in activity patterns between weekdays and weekends?

```{r}
wee_sum <- function(activity) {
  workweek <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
  activity$workday <- weekdays(activity$date) %in% workweek
  gro_by_wee_day <- group_by(activity, interval, workday)
  summary_tbl <- summarise(gro_by_wee_day, average_steps = mean(steps, na.rm=TRUE))
  return(summary_tbl)
}

plot_it <- function(wee_dif) {
  par(mfrow = c(2,1))
  with(wee_dif[wee_dif$workday==FALSE,],
       plot(interval, average_steps, type="l", main="Weekend",
            xlab="Interval", ylab="Steps"))
  with(wee_dif[wee_dif$workday==TRUE,],
       plot(interval, average_steps, type="l", main="Weekday",
            xlab="Interval", ylab="Steps"))
}

wee_dif <- wee_sum(activity)
plot_it(wee_dif)
```